\pkg{cubble}: An R Package for Organizing and Wrangling Multivariate Spatio-temporal Data

### 2022-06-14 4-3-water-level.Rmd

```{r}
climate <- weatherdata::climate_full |>
  filter(between(stringr::str_sub(id, 7, 8), 76, 90)) |>
  face_temporal() |>
  filter(lubridate::year(date) == 2020) |>
  face_spatial() |>
  mutate(type = "climate")

@@

```{r echo = TRUE}
res <- match_sites(
  river, climate,
  temporal_by = c("Water_course_level" = "prcp"),
  temporal_independent = "prcp",  
  temporal_n_highest = 30,
  temporal_min_match = 15
)
```

@@

```{r}
res |> print(n = 8)
```

>>>>

```{r eval = FALSE}
climate <- weatherdata::climate_full |>
  filter(between(stringr::str_sub(id, 7, 8), 76, 90)) |>
  face_temporal() |>
  filter(lubridate::year(date) == 2020) |>
  face_spatial() |>
  mutate(type = "climate")

@@

```{r echo = TRUE,  eval = FALSE}
res <- match_sites(
  river, climate,
  temporal_by = c("Water_course_level" = "prcp"),
  temporal_independent = "prcp",  
  temporal_n_highest = 30,
  temporal_min_match = 15
)
```

@@

```{r,  eval = FALSE}
res |> print(n = 8)
```

### 2022-06-14 jss/cubble-jss.tex

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_unfold |> face_spatial() |> face_temporal()
\end{CodeInput}
\begin{CodeOutput}
# cubble:  date, id [5]: long form
# bbox:    [115.97, -32.94, 133.55, -12.42]
# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id
#   [dbl]
  id          date        prcp  tmax  tmin
  <chr>       <date>     <dbl> <dbl> <dbl>
1 ASN00009021 2020-01-01     0  31.9  15.3
2 ASN00009021 2020-01-02     0  24.9  16.4
3 ASN00009021 2020-01-03     6  23.2  13  
4 ASN00009021 2020-01-04     0  28.4  12.4
5 ASN00009021 2020-01-05     0  35.3  11.6
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cb <- as_cubble(
+   list(spatial = lga, temporal = covid),
+   key = lga, index = date, coords = c(cent_long, cent_lat)
+   )
\end{CodeInput}
\begin{CodeOutput}
! Some sites in the temporal table don't have spatial information
\end{CodeOutput}
\begin{CodeOutput}
! Some sites in the spatial table don't have temporal information
\end{CodeOutput}
\begin{CodeOutput}
! Use argument `output = "unmatch"` to check on the unmatched key
\end{CodeOutput}
\end{CodeChunk}

@@

The output from matching is also a cubble, with additional column \code{dist} and \code{group} produced from spatial matching and \code{n_match} from temporal matching.

Figure \ref{fig:matching} plots the matched pairs on the map. There are four pairs of matches, all located in the middle of Victoria and the expected concurrent increase in precipitation and water level can be observed.

>>>>

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_unfold |> face_spatial() |> face_temporal()
\end{CodeInput}
\begin{CodeOutput}
# cubble:  date, id [5]: long form
# bbox:    [115.97, -32.94, 133.55, -12.42]
# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id
#   [dbl]
  id          date        prcp  tmax  tmin  long   lat
  <chr>       <date>     <dbl> <dbl> <dbl> <dbl> <dbl>
1 ASN00009021 2020-01-01     0  31.9  15.3  116. -31.9
2 ASN00009021 2020-01-02     0  24.9  16.4  116. -31.9
3 ASN00009021 2020-01-03     6  23.2  13    116. -31.9
4 ASN00009021 2020-01-04     0  28.4  12.4  116. -31.9
5 ASN00009021 2020-01-05     0  35.3  11.6  116. -31.9
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cb <- as_cubble(
+   list(spatial = lga, temporal = covid),
+   key = lga, index = date, coords = c(cent_long, cent_lat)
+   )
\end{CodeInput}
\begin{CodeOutput}
! Some sites in the temporal table don't have spatial information
\end{CodeOutput}
\begin{CodeOutput}
! Some sites in the spatial table don't have temporal information
\end{CodeOutput}
\begin{CodeOutput}
! Use argument `output = "unmatch"` to check on the unmatched key
\end{CodeOutput}
\begin{CodeOutput}
Joining, by = "lga"
\end{CodeOutput}
\end{CodeChunk}

@@

The output from this function is also a cubble, with additional columns \code{dist} and \code{group} produced from spatial matching, and \code{n_match} from temporal matching.

Figure \ref{fig:matching} (A) shows the four matched pairs on the map. For these the expected concurrent increase in precipitation and water level can be seen clearly when comparing these variables on a standardised scale, see Figure \ref{fig:matching} (B).
